---
title: "Statistical Consulting"
subtitle: "Data Preprocessing"
author: "Dennis Klein, Daniel Seussler "
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  cleanrmd::html_document_clean:
    keep_md: true
    theme: tacit  # see cleanrmd::cleanrmd_themes()
                  # or set to NULL for dynamic theme picker (requires internet)
---

# Start

## Sources

* CEPII TRADEHIST <http://www.cepii.fr/cepii/en/bdd_modele/presentation.asp?id=32>
    * Bilateral Trade and Bilateral Tariffs
* Exchange Rates <https://www.measuringworth.com/datasets/exchangeglobal/>
* CPI US <https://www.statbureau.org/en/united-states/cpi-u>
* Replication files <https://www.cambridge.org/core/journals/network-science/article/separable-and-semiparametric-networkbased-counting-processes-applied-to-the-international-combat-aircraft-trades/0D57EC7B7E1775B0BEF72BDE101E507F#article>

## ToDo's

* Alternative Handelsdaten vom Ifo Institut?
* Rebase nominal trade values for constant USD.


## Setup

```{r message = FALSE}
library(countrycode)
library(readxl)
library(tidyverse)
```


# Bilateral Tradeflows

## UN COMTRADE

Useful extension for 2014 onwards?

## TRADEHIST

Load trade data. 

```{r}
TRADHIST_BITRADE_BITARIFF_1 <- read_excel("data raw/CEPII TRADEHIST/TRADHIST_BITRADE_BITARIFF_1.xlsx")
TRADHIST_BITRADE_BITARIFF_2 <- read_excel("data raw/CEPII TRADEHIST/TRADHIST_BITRADE_BITARIFF_2.xlsx")
TRADHIST_BITRADE_BITARIFF_3 <- read_excel("data raw/CEPII TRADEHIST/TRADHIST_BITRADE_BITARIFF_3.xlsx")
```

Currency British Pound Sterling. Nominal trade flows, rebase to constant 2000$ USD. Or something else?

```{r}
exchange_usd_gbp <- read.csv(file = "data raw/EXCHANGEGLOBAL_1945-2020.csv", 
                             header = FALSE) %>%
  rename(year = V1, GBP_USD = V2) %>%
  mutate(GBP_USD = as.numeric(gsub(" British Pound", "", GBP_USD)))

cpi_us <- read.csv(file = "data raw/united-states.index.cpi-u (statbureau.org).csv") %>%
  rename(year = ï..Year) %>%
  filter(year %in% c(1950:2020)) %>%
  pivot_longer(cols = c("January", "February", "March", "April", "May", "June",
                        "July", "August", "September", "October", "November",
                        "December"),
               names_to = "month", values_to = "cpi") %>% 
  group_by(year) %>%
  summarize(cpi = mean(cpi)) %>%
  mutate(cpi_factor = cpi / as.numeric(filter(., year == 2000)[,2]))
```

Merge data, filter for out time-frame, 1950 and newer (?). Do conversions. Replace flow_0 values. 

```{r}
trade <- TRADHIST_BITRADE_BITARIFF_1 %>%
  bind_rows(TRADHIST_BITRADE_BITARIFF_2) %>%
  bind_rows(TRADHIST_BITRADE_BITARIFF_3) %>%
  filter(year %in% c(1950:2020)) %>%
  mutate(FLOW = replace(FLOW, FLOW_0 == 0, 0)) %>%
  left_join(exchange_usd_gbp, by = c("year" = "year"), na_matches = "never") %>%
  left_join(cpi_us, by = c("year" = "year"), na_matches = "never") %>%
  mutate(FLOW = (FLOW / GBP_USD) / cpi_factor) %>% 
  select(iso_o, iso_d, year, FLOW)
```

Countries included in our time frame. Useful Coverage? Selection?

```{r}
countries <- data.frame(iso = c(trade$iso_o, trade$iso_d)) %>% 
  unique() %>%
  mutate(name = countrycode(iso, "iso3c", "country.name")) %>%
  arrange(name)

head(countries)
```

# MCW Weapon Transfers

Replication files from *Separable and semiparametric network-based counting processes applied to the international combat aircraft trades*.




